
```{r}
library(ISLR)
library(tidyverse)
library(tidyselect)
library(pastecs)
library(tibble)
library(data.table)
library(ggplot2)
library(pwr)
library(WebPower)
library(stringr)
#library(WebPower/lavaan)
```


```{r}
df <- read.csv(".\\dataset\\mortgage.csv")
##-----EDA process
head(df, 5)
str(df)
```
```{r}
#descriptive analysis
summary(df)
```

```{r}
#remove col of id
df<-select(df,-c("id"))
head(df, 5)
#check for missing values
sum(is.na(df))
```
```{r}
df <- df[filter <- !is.na(df$LTV_time),]
dim(df)
```

```{r}
boxplot(df)
```
It is obviously that there are outliers in the dataset, which may affect the model.

```{r}
#select numerical features
num_var = c("balance_time", "LTV_time", "interest_rate_time", "hpi_time", "gdp_time", "uer_time", "balance_orig_time", "FICO_orig_time", "LTV_orig_time", "Interest_Rate_orig_time","hpi_orig_time")

df_numerical<- df %>% select_if(names(.) %in% num_var)
str(df_numerical)

```
https://stackoverflow.com/questions/4877357/how-to-plot-all-the-columns-of-a-data-frame-in-r  
https://www.statisticssolutions.com/univariate-and-multivariate-outliers/#:~:text=A%20multivariate%20outlier%20is%20a,data%20to%20contain%20extreme%20cases.
It is obviously that there are outliers in the dataset, which may affect the model. 
balance_time, mat_time, LTV_orig_time, interest_rate_orig_time, hpi_orig_time, gdp_time
```{r}
par(mfrow=c(4,3),mar=c(0.7, 3, 2, 1.5)) 
len_num <- length(names(df_numerical))
for(i in 1:len_num){
  mname <- substitute(df_numerical[,i])
  boxplot(df_numerical[,i],main=names(df_numerical[i]))
    }
#dfplot
```
```{r}
#plot categorical features
`%!in%` <- Negate(`%in%`)
df_categorical<- df %>% select_if(names(.) %!in% num_var)
#str(df_categorical)
par(mfrow=c(4,3),mar=c(2, 3, 2.5, 2)) 
len_cat <- length(names(df_categorical))
for(i in 1:len_cat){
  mname <- substitute(df_categorical[,i])
  hist(df_categorical[,i],main=names(df_categorical[i]))
    }
```


```{r}
#detect outlier
outliers_balance_time = boxplot.stats(df$balance_time)
outliers_LTV = boxplot.stats(df$LTV_time)

#outliers
#detect >19000 outlier values out of 622489 
```

```{r}
##Detect with z-score
#calculate the score
#z_scores <- as.data.frame(sapply(df, function(df) (abs(df - mean(df))/sd(df))))
#head(z_scores)
```
Z-scores can quantify the unusualness of an observation when your data follow the normal distribution. The further away an observation’s Z-score is from zero, the more unusual it is. A standard cut-off value for finding outliers are Z-scores of +/-3 or further from zero. 
https://statisticsbyjim.com/basics/outliers/#:~:text=Using%20Z%2Dscores%20to%20Detect%20Outliers&text=A%20Z%2Dscore%20of%20zero,value%20that%20equals%20the%20mean.&text=The%20further%20away%20an%20observation's,3%20or%20further%20from%20zero .
```{r}
#function calculating z-score of a column
z_score_cal <-function (col){
  zcol <- abs(col - mean(col))/sd(col)
}
```

```{r}
df_1 <- df
#len_df = length(names(df_1))

##-----remove values whose z-score < 3

#df_1 %>% select(names(.) %in% num_var) 
#  df_1[z_score_cal(names(.)) < 3,]
    
#balance_time,  LTV_time, interest_rate_time, balance_orig_time, FICO_orig_time, LTV_orig_time, interest_rate_orig_time, hpi_orig_time. 

df_1 <- df_1[z_score_cal(df_1$balance_time) < 3,]
df_1 <- df_1[z_score_cal(df_1$LTV_time) < 3,]
df_1 <- df_1[z_score_cal(df_1$interest_rate_time) < 3,]
df_1 <- df_1[z_score_cal(df_1$FICO_orig_time) < 3,]
df_1 <- df_1[z_score_cal(df_1$LTV_orig_time)< 3, ]
df_1 <- df_1[z_score_cal(df_1$balance_orig_time)<3,]
df_1 <- df_1[z_score_cal(df_1$Interest_Rate_orig_time)<3,]
df_1 <- df_1[z_score_cal(df_1$hpi_orig_time)<3,]
dim(df_1)

```

```{r}
#plot to see the distribution after outlier removal
boxplot(df_1)
```

Calculate sample size 
```{r}
prop_status = prop.table(table(df$status_time))
(prop_status)
```

```{r}
#chi = chisq.test(df$status_time)
 wp.logistic(p0 = 0.043, p1 = 0.0024, alpha = 0.05, power = 0.8, alternative="two.sided", family="normal")
```

```{r}
 pwr.p.test(h=0.2, sig.level=0.05, power=0.80, alternative="two.sided")
```


